Measuring matter effects in Neutron star - Black hole binaries with Advanced LIGO

### 2016-04-07 Paper/paper.tex
minor change to bayesian sec

Inferred from $d_n$, the joint probability distribution for binary parameters
$\vec{\Theta}:=\vec{\theta}\cup\{\lambdans\}$ is given by
\begin{equation}\label{eq:postprob}
 p(\vec{\Theta} | d_n, K) = \dfrac{p(d_n|\vec{\Theta}, K)\,p(\vec{\Theta} | K)}{p(d_n|K)}.
\end{equation}
Here, $p(\vec{\Theta} | K)$ is the prior probability for binary parameters of
taking particular values. In this paper, we use a uniform prior on the
individual masses within pre-chosen ranges, as well as a uniform prior on the
spin of the black hole. We set neutron star's spin $=0$, and its tidal
deformability parameter $\lambdans$ is sampled uniformly from $[0, 4000]$. In
addition, we exclude mass-ratios between $[1,2]$ from our prior. This is done
because of model constraints for LEA+. The prior probability of obtaining a 
particular realization of data $p(d_n|K)$ is absorbed into the overall 
normalization. Finally, the first term in the numerator
$p(d_n|\vec{\theta}, \lambdans, K)$ is the likelihood of obtaining the given
stretch of data $d_n$ if we assume that a signal parameterized by
$\vec{\theta}\cup\{\lambdans\}$ is buried in it, i.e.
\begin{equation}\label{eq:likelihood}
 \LL(\vec{\Theta}) \equiv p(d_n| \vec{\Theta}, K) = \mathcal{N} \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
\end{equation}
where $h\equiv h(\vec{\Theta})$ is a template given by our waveform model LEA+,
and the detector noise weighted inner-product $\langle a|b\rangle$ is defined 
as
\begin{equation}
\langle a|b\rangle \equiv 4\,\mathrm{Re}\left[\int_0^\infty \dfrac{\tilde{a}(f) \tilde(b)(f)^*}{S_n(|f|)}\,\D f\right],
\end{equation}
% 
where $\tilde{a}(f)$ is the Fourier transform of the finite time series $a(t)$,
and $S_n(|f|)$ is the one-sided amplitude spectrum of detector noise. The 
definition of $\LL$ in Eq.~\ref{eq:likelihood} assumes that the instrument
noise is colored Gaussian.

We sample the probability distribution given in Eq.~\ref{eq:postprob} using
an ensemble sampler Markov-chain Monte-Carlo algorithm based on the {\tt emcee}
package~. A total of $100$ independent walkers were used, each of which
was allowed to collect $100,000$ samples. We measure the correlation time and keep
only $20,000$ independent posterior points to further analyze the calculated 
posterior probability distributions. One simplifying assumption made to mitigate 
computational cost was setting the frequency sampling rate $\Delta f=0.4$~Hz, which
we found to be sufficient to calculate likelihoods robustly~\cite{Purrer:2015nkh}.
\prayush{TODO: MICHAEL}.

>>>>

Inferred from $d_n$, the joint probability distribution for binary parameters
$\vec{\Theta}:=\vec{\theta}\cup\{\lambdans\}$ is given by
\begin{equation}\label{eq:postprob}
 p(\vec{\Theta} | d_n, K) = \dfrac{p(d_n|\vec{\Theta}, K)\,p(\vec{\Theta} | K)}{p(d_n|K)}.
\end{equation}
Here, $p(\vec{\Theta} | K)$ is the prior probability for binary parameters of
taking particular values. In this paper, we use a uniform prior on the
individual masses within pre-chosen ranges, as well as a uniform prior on the
spin of the black hole. We set neutron star's spin $=0$, and its tidal
deformability parameter $\lambdans$ is sampled uniformly from $[0, 4000]$. In
addition, we exclude mass-ratios between $[1,2]$ from our prior. This is done
because of model constraints for LEA+. The prior probability of obtaining a 
particular realization of data $p(d_n|K)$ is absorbed into the overall 
normalization. Finally, the first term in the numerator
$p(d_n|\vec{\theta}, \lambdans, K)$ is the likelihood of obtaining the given
stretch of data $d_n$ if we assume that a signal parameterized by
$\vec{\theta}\cup\{\lambdans\}$ is buried in it, i.e.
\begin{equation}\label{eq:likelihood}
 \LL(\vec{\Theta}) \equiv p(d_n| \vec{\Theta}, K) = \mathcal{N} \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
\end{equation}
where $h\equiv h(\vec{\Theta})$ is a template given by our waveform model LEA+,
and the detector noise weighted inner-product $\langle a|b\rangle$ is defined 
as
\begin{equation}
\langle a|b\rangle \equiv 4\,\mathrm{Re}\left[\int_0^\infty \dfrac{\tilde{a}(f) \tilde(b)(f)^*}{S_n(|f|)}\,\D f\right],
\end{equation}
% 
where $\tilde{a}(f)$ is the Fourier transform of the finite time series $a(t)$,
and $S_n(|f|)$ is the one-sided amplitude spectrum of detector noise. The 
definition of $\LL$ in Eq.~\ref{eq:likelihood} assumes that the instrument
noise is colored Gaussian. The marginalized probability distribution for any
single binary parameter (say $\alpha$) can be obtained by integrating
Eq.~\ref{eq:postprob} over all other parameters, i.e. 
\begin{equation}
 p(\alpha | d_n, K) = \int\D \vec{\Theta}_\alpha\, p(\vec{\Theta} | d_n, K),
\end{equation}
where $\vec{\Theta}_\alpha$ is the set of remaining parameters, i.e.
$\vec{\Theta}_\alpha:=\vec{\Theta} - \{\alpha\}$.

We sample the probability distribution of Eq.~\ref{eq:postprob} using
an ensemble sampler Markov-chain Monte-Carlo algorithm based on the {\tt emcee}
package~. A total of $100$ independent walkers were used, each of which
was allowed to collect $100,000$ samples. We measure the correlation time and keep
only $20,000$ independent posterior points to further analyze the calculated 
posterior probability distributions. One simplifying assumption made to mitigate 
computational cost was setting the frequency sampling rate $\Delta f=0.4$~Hz, which
we found to be sufficient to calculate likelihoods robustly~\cite{Purrer:2015nkh}.
\prayush{TODO: MICHAEL}.
